---
output: github_document
---

<!-- README.md is generated from README.Rmd — edit that file, then rmarkdown::render("README.Rmd"). -->

# piknit

<!-- badges: start -->
[![License: GPL v2+](https://img.shields.io/badge/License-GPL%20v2%2B-blue.svg)](https://www.gnu.org/licenses/gpl-2.0.html)
<!-- badges: end -->

**A way to embed the [Pi](https://github.com/earendil-works) coding agent in R.** Drive a live agent from R with a
small scripting API — `pi_run()` for one turn, `pi_session()` for a memory-keeping multi-turn session — or drop it
straight into R Markdown / Quarto through knitr engines that run a *live* agent (and fail-loud shell commands) at
render time, weaving the agent's real replies and command output into the document. **Rendering becomes an
integration test**: a stale claim or a broken command fails the build. Same idea as `leanknit` and `duckknit`,
pointed at an agent.

Two ways to use it:

- **As an R binding** — `pi_run()` / `pi_session()` call a Pi agent from any R script, app, or pipeline; the agent
  (with its tools and extensions) becomes a function you call and whose reply you keep.
- **As literate engines** — ```` ```{pi} ```` and ```` ```{pish} ```` chunks put a live agent and fail-loud
  commands into a rendered document.

```{r setup, include = FALSE}
library(piknit)
knitr::opts_chunk$set(comment = "")
# the live chunks need the pi CLI; where it is absent (e.g. CI) they are skipped
pi_ok <- nzchar(Sys.which("pi"))
```

## Install

```r
# remotes::install_github("sounkou-bioinfo/piknit")
library(piknit)
```

Needs the `pi` CLI on the `PATH`. Set defaults with `options(piknit.model=, piknit.provider=)` or the
`PIKNIT_PI` env var.

## Call the agent from R

The agent is a function you call; its reply is a character vector you keep — run live when this README renders:

```{r call-from-r, eval = pi_ok}
reply <- pi_run("In one sentence, what is a coding agent?")
reply
```

Point it at an extension (`extension = "path/to/index.ts"`) to give the agent tools. That is the whole embedding:
`pi_run()` is one turn, `pi_session()` is a memory-keeping conversation, and the knitr engines below are just these
calls wired into a render.

## The `pi` engine — the chunk body is the prompt

Write a prompt in a ```` ```{pi} ```` chunk; at render time a live agent turn runs and its reply is woven in under
the exact command used. A small, fast model (`gpt-5.3-codex-spark` by default) keeps a live turn to seconds.

````
```{pi}`r ''`
In one sentence, what is literate programming?
```
````

```{pi}
In one sentence, what is literate programming?
```

## The `pish` engine — fail-loud shell

Run a command; a non-zero exit **stops the render** (never bakes a failure as an answer), and JSON output is
fenced for highlighting.

````
```{pish}`r ''`
node --version
```
````

## Set the project root

pi treats its working directory as the **project root** — what scopes the agent's file tools and where relative
extension / skill / manifest paths resolve. Point any call at a root with `dir`:

```r
pi_run(
  "Count variants by consequence in examples/variant-counts/manifest.json.",
  dir = "~/pi-bio-agent",
  extension = "extensions/pi-coding-agent/index.ts"
)
```

## Stream the response

Because pi's `--mode rpc` emits JSON events as they happen, `pi_stream()` delivers the reply incrementally —
assistant text as deltas, every event to a callback — ending on pi's `agent_end`:

```r
pi_stream("Explain content-addressed storage in two sentences.")          # prints as it is typed
pi_stream("...", on_event = function(ev) if (ev$type == "turn_end") message("done"))
```

## A persistent session across chunks

`pi_session()` gives successive turns shared conversation memory (via a stable `--session-id`) without holding a
long-lived process, so an earlier chunk's result is in scope later:

```r
s <- pi_session(dir = "~/pi-bio-agent", extension = "extensions/pi-coding-agent/index.ts")
s$ask("Load examples/variant-counts/manifest.json and count variants by consequence.")
s$ask("Now express the same counts as percentages.")  # remembers the table above
```

## Why

Docs that *run* don't drift. Pointing the render at a live agent turns a README, a vignette, or a tutorial into a
real end-to-end test of the agent, its tools, and the claims you make about them — the discipline that caught
real bugs while building [pi-bio-agent](https://github.com/sounkou-bioinfo/pi-bio-agent), piknit's first user.

## License

GPL (>= 2) © sounkou-bioinfo
